---
title: " TA Section 2 Code Exercise"
author: "150B/355B Introduction to Machine Learning"
date: "1/19/2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Question 1 Time for Change Model

## 1. Setting up the time for a change model

Download and create in object from the .csv file ``TimeChange.csv" into your computer. 

```{r}
rm(list=ls())
#Change to your directory for this section
setwd('~/Dropbox (IPL)/150B Machine Learning/02_Section') 
d<-read.csv("TimeChange.csv")

#Bivariate model of VoteShare on Incumbency Approval
bivariate <- lm(IncumbentVoteShare~Incumbent_Net_Approval, data = d)
bivariate$coefficients
```

## 2. Fitting Multivariate Regression

```{r}
bivariate <- lm(IncumbentVoteShare~Incumbent_Net_Approval, data = d)
multivariate <- lm(IncumbentVoteShare~Incumbent_Net_Approval + Q1_GDP_Growth + 
                     Q2_GDP_Growth + Incumbent_Party_Two_Terms, data = d)

summary(bivariate)
summary(multivariate)
```


## 3. Multivariate Prediction

### 3.1

Write a prediction function using the output from the multivariate regression.
```{r}
# YOUR CODE HERE
```

### 3.2 

Compare it to fitted values.
```{r}
# YOUR CODE HERE
```


## 4. Linear Algebra in R ###

### 4.1

Create three vectors in R.  Make two of the vectors of length 4 and one vector of length 5.
- vector 1 should equal (1, 2, 3, 4)
- vector 2 should equal (3, 4, 5, 6)
- vector 3 should equal (7, 8, 9, 10, 11)

```{r}
# YOUR CODE HERE
```
### 4.2

Confirm the length of the vectors using the length function.

```{r}
# YOUR CODE HERE
```

### 4.3

Now, we're going to use `%*%` to take their inner product. First, take the inner product of vector 1 and vector 2. Second, write the inner product of vector 1 and vector 3. What do you notice? 

```{r}
# YOUR CODE HERE
```

## 5. Vector based regression

### 5.1 

Write a function to predict from a multivariate regression using vectors.

```{r}
# YOUR CODE HERE
```

### 5.2

Compare the predicted values between you two predict functions (one with linear algebra, one without) and from the lm ("canned") model.
```{r}
# YOUR CODE HERE
```


## Question 2 Predicting Iraq Vote using LPM

We're going to analyze the Senate's vote on the Iraq war authorization.We'll find this data set in the PSCL package, which is called `pscl` in R.

To get started, we need to install and load the library. Execute the following code to install pscl.

```{r}
rm(list=ls())
setwd('~/Dropbox (IPL)/150B Machine Learning/02_Section')
#install.package('pscl')

library(pscl)
data(iraqVote)
```

Checking the column names of the data set we have:
- y : The vote on the Iraq war authorization
- state.abb : Name of the senator's state
- name : senator's name
- rep : an indicator = TRUE if the senator is a Republican, FALSE if not a Republican
- state.name : the name of the senator's state
- gorevote: the share of the two party vote cast for Al Gore in the 2000 election


### 1 
Let's fit a linear probability model. Using a linear regression, regress the vote against `rep` and `gorevote`.

```{r}
fit<-lm(y~rep+gorevote, data=iraqVote) #fit linear model and print result
summary(fit)
fit$coefficients #extract the vector of coefficients
```

### 2

Using the regression, calculate predicted probabilities for each of the observations.  You can do this by creating an appropriate matrix or with a for loop.

```{r}
# create matrix of predictors
indep.matrix <- cbind(1, iraqVote$rep, iraqVote$gorevote)

# multiply by coefficients
pred_prob_lm<- c(indep.matrix%*%fit$coef) 

# we can also take the fitted values from the model
fitted_lm <- fit$fitted.values

# Are they the same?
pred_prob_lm[1:10]
fitted_lm [1:10]
```

### 3

Summarize the predicted probabilities. What do you notice about the predicted probabilities? (in particular, the maximum and minimum values)
```{r}
# YOUR CODE HERE
```

### 4

Write a function that takes the predicted probabilities and a threshold, and returns a classification decision. Use that function to classify each senator as Yay or Nay on the Iraq vote.
```{r}
# YOUR CODE HERE
```

### 5

What proportion of senators do you classify as a likely Iraq Vote? What proportion of Democrats and Republicans?
```{r}
# YOUR CODE HERE
```