---
title: "Interactive Code Lecture 2"
author: "150B/355B Introduction to Machine Learning"
date: "1/11/2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## 1. Setting up the time for a change model

### 1.1

Download and create in object from the .csv file ``TimeChange.csv" into your computer. What class is the object you just created?

```{r}
## rm(list=ls())
setwd('/Users/tongtongzhang/Dropbox (IPL)/155B Machine Learning/Lecture/Lecture 2')
dat <- read.csv("TimeChange.csv")
class(dat)
```

### 1.2

Print the dimensions of the matrix and the column names.

```{r}
dim(dat)
names(dat)
```

### 1.3

Provide Summary Statistics for each variable in the data set.

```{r}
summary(dat)
```

STOP!


## 2. Simple Regression

### 2.1 

Our first endeavor will be to predict IncumbentVoteShare with Incumbent_Net_Approval.

Create a plot of IncumbentVoteShare against Incumbent_Net_Approval. Try using the `identify` function to find interesting observations.

```{r}
plot(dat$Incumbent_Net_Approval, dat$IncumbentVoteShare, pch = 19, xlab = "Net Approval", 
     ylab = "Incumbent Vote Share") 

identify(dat$Incumbent_Net_Approval, dat$IncumbentVoteShare, labels = dat$Year)
```

### 2.2

Create an object that contains the regression of `IncumbentVoteShare` against `Incumbent_Net_Approval`.

```{r}
lin <- lm(IncumbentVoteShare ~ Incumbent_Net_Approval, data = dat)
```

### 2.3

Include a regression line of `IncumbentVoteShare` against `Incumbent_Net_Approval` on the plot (hint: abline might be useful.)

```{r}
plot(dat$IncumbentVoteShare ~ dat$Incumbent_Net_Approval, pch = 19, xlab = "Net Approval", 
     ylab = "Incumbent Vote Share")
abline(lin, col = "red")
```

### 2.4 

Now, using the regression object you used to add the line to your plot, check the names that are provided in the `lm` object. Summarize:
1) Fitted Values
2) Residuals

And store the coefficients in a vector.

```{r}
names(lin)
lin$coefficients[2]
```

```{r}

summary(lin$fitted.values)
# What are the fitted values?

```

```{r}

summary(lin$residuals)
# What are the residuals?
```
STOP!




## 3. Prediction 

### 3.1

`predict` is an extremely useful function that you've likely worked with before. Today, we're going to write our own function using an `lm` object.

Implement a prediction function that takes as an argument an lm object and a value for net approval.

```{r}
my_pred <- function(mod, variable){
  coefs <- mod$coefficients
  pred <- coefs[1] + coefs[2]*variable
  return(pred)
}
```

### 3.2 

Report the prediction for the following values: -4, 100, -100, 50
 
```{r}

new_obs <- c(-4, 100, -100, 50) # what do these values substantively mean
my_pred(lin, new_obs) # what is this returning?

predict(lin, newdata=data.frame(Incumbent_Net_Approval=new_obs))
```

